### 2025-03-10 paper/preprint/paper.tex
Add shortened discussion of hyper-parameters to the body

@@

@@

>>>>

We run the experiments with three CE generators: \emph{Generic} of
Wachter, Mittelstadt, and Russell
(\citeproc{ref-wachter2017counterfactual}{2017}) as a simple baseline
approach, \emph{REVISE} (\citeproc{ref-joshi2019realistic}{Joshi et al.
2019}) that aims to generate plausible counterfactuals using a surrogate
Variational Autoencoder (VAE), and \emph{ECCo}---the generator of
Altmeyer et al. (\citeproc{ref-altmeyer2023faithful}{2023}) but without
the conformal prediction component---as a method that directly targets
both faithfulness and plausibility of the CEs.

@@

\subsubsection{Plausibility}\label{plausibility-1}

\subsubsection{Actionability}\label{actionability-1}

\subsubsection{Impact of hyperparameter
settings}\label{impact-of-hyperparameter-settings}

We extensively test the impact of three types of hyperparameters on the
proposed training regime. Our complete results are available in the
technical appendix; this section focuses on the main findings.

\textbf{\emph{Hyperparameters of the CE generators.}} First, we observe
that CT is highly sensitive to hyperparameter settings but (a) there are
manageable patterns and (b) we can typically identify settings that
improve either plausibility or cost, and commonly both of them at the
same time. Second, we note that the choice of a CE generator has a major
impact on the results. For example, \emph{REVISE} tends to perform the
worst, most likely because it attempts to explain the decisions of a VAE
that is an imperfect surrogate of the original model. Third, increasing
\(T\), the maximum number of steps, generally yields better outcomes
because more CEs can mature in any iteration. Fourth, the impact of
\(\tau\), the required decision threshold is more difficult to predict.
On ``harder'' datasets it may be difficult to satisfy high \(\tau\) for
any given sample (i.e., also factuals) and so increasing this threshold
does not seem to correlate with better outcomes.

\textbf{\emph{Hyperparameters for penalties}} We find that the strength
of the energy regularization, \(\lambda_{\text{reg}}\) is highly
impactful; energy must be sufficiently regularized to avoid poor
performance in terms of decreased plausibility and increased costs. The
sensitivity with respect to \(\lambda_{\text{div}}\) and
\(\lambda_{\text{adv}}\) is much less evident. While high values of
\(\lambda_{\text{reg}}\) may increase the variability in outcomes when
combined with high values for Equation~\ref{eq-obj} penalties, this
effect is not very pronounced.

\textbf{\emph{Other hyperparameters.}} We observe that the effectiveness
and stability of CT is positively associated with the number of
counterfactuals generated during each training epoch. We also confirm
that a higher number of training epochs is beneficial. Interestingly, we
find that it is not necessary to employ CT during the entire training
phase to achieve the desired improvements in explainability. We have
tested training models conventionally during the first half of training
before switching to CT after this initial ``burn-in'' period and
observed positive results. Put differently, CT may be a way to improve
the explainability of trained models in a fine-tuning manner.

@@

\bibitem[\citeproctext]{ref-altmeyer2023faithful}
Altmeyer, Patrick, Mojtaba Farmanbar, Arie van Deursen, and Cynthia C.
S. Liem. 2023. {``Faithful Model Explanations Through Energy-Constrained
Conformal Counterfactuals.''} \url{https://arxiv.org/abs/2312.10648}.

### 2025-03-10 paper/sections/experiments.qmd
Add shortened discussion of hyper-parameters to the body

@@

>>>>

We run the experiments with three CE generators: *Generic* of @wachter2017counterfactual as a simple baseline approach, *REVISE* [@joshi2019realistic] that aims to generate plausible counterfactuals using a surrogate Variational Autoencoder (VAE), and *ECCo*---the generator of @altmeyer2023faithful but without the conformal prediction component---as a method that directly targets both faithfulness and plausibility of the CEs.

@@

### Plausibility

### Actionability

### Impact of hyperparameter settings
We extensively test the impact of three types of hyperparameters on the proposed training regime. Our complete results are available in the technical appendix; this section focuses on the main findings.

***Hyperparameters of the CE generators.*** First, we observe that CT is highly sensitive to hyperparameter settings but (a) there are manageable patterns and (b) we can typically identify settings that improve either plausibility or cost, and commonly both of them at the same time. Second, we note that the choice of a CE generator has a major impact on the results. For example, *REVISE* tends to perform the worst, most likely because it attempts to explain the decisions of a VAE that is an imperfect surrogate of the original model. Third, increasing $T$, the maximum number of steps, generally yields better outcomes because more CEs can mature in any iteration. Fourth, the impact of $\tau$, the required decision threshold is more difficult to predict. On "harder" datasets it may be difficult to satisfy high $\tau$ for any given sample (i.e., also factuals) and so increasing this threshold does not seem to correlate with better outcomes.

***Hyperparameters for penalties*** We find that the strength of the energy regularization, $\lambda_{\text{reg}}$ is highly impactful; energy must be sufficiently regularized to avoid poor performance in terms of decreased plausibility and increased costs. The sensitivity with respect to $\lambda_{\text{div}}$ and $\lambda_{\text{adv}}$ is much less evident. While high values of $\lambda_{\text{reg}}$ may increase the variability in outcomes when combined with high values for @eq-obj penalties, this effect is not very pronounced.

***Other hyperparameters.*** We observe that the effectiveness and stability of CT is positively associated with the number of counterfactuals generated during each training epoch. We also confirm that a higher number of training epochs is beneficial. Interestingly, we find that it is not necessary to employ CT during the entire training phase to achieve the desired improvements in explainability. We have tested training models conventionally during the first half of training before switching to CT after this initial "burn-in" period and observed positive results. Put differently, CT may be a way to improve the explainability of trained models in a fine-tuning manner.
